---
title: "Strings and Factors"
author: "My An Huynh"
date: "2024-10-10"
output: github_document
---
```{r setup, include=FALSE}
library(tidyverse)
library(rvest)
library(p8105.datasets)

```

String matching needs to be the exact match. 
```{r}
string_vec = c("My", "name", "is", "Claudia")

str_detect(string_vec, "a")

str_replace(string_vec, "Claudia", "claudia")

str_replace(string_vec, "e", "E")

```

Use ^ to denote the beginning of the string and $ to denote the end of the string.

```{r}
string_vec = c(
  "i think we all rule for participating",
  "i think i have been caught",
  "i think this will be quite fun actually",
  "it will be fun, i think"
  )

str_detect(string_vec, "^i think")

str_detect(string_vec, "i think$")
```

```{r}
string_vec = c(
  "Time for a Pumpkin Spice Latte!",
  "went to the #pumpkinpatch last weekend",
  "Pumpkin Pie is obviously the best pie",
  "SMASHING PUMPKINS -- LIVE IN CONCERT!!"
  )

str_detect(string_vec, "pumpkin")
str_detect(string_vec,"[Pp]umpkin")
```

```{r}
string_vec = c(
  '7th inning stretch',
  '1st half soon to begin. Texas won the toss.',
  'she is 5 feet 4 inches tall',
  '3AM - cant sleep :('
  )

str_detect(string_vec, "[0-9][a-zA-Z]")
```

## Factors

```{r}
sex_vec = factor(c("male", "male", "female", "female"))
as.numeric(sex_vec)

sex_vec = fct_relevel(sex_vec, "male")
```

## Revisit lots of examples

```{r}
url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

drug_use_html = 
  read_html(url)

marj_use_df = 
  drug_use_html |> 
  html_table() |> 
  first() |>
  slice(-1) |> 
  select(-contains("P value")) |> 
  pivot_longer(
    cols = -State, 
    names_to = "age_year",
    values_to = "percent"
  ) |> 
  separate(age_year, into = c("age", "year"), sep = "\\(") |> 
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_remove(percent, "[a-c]$"),
    percent = as.numeric(percent)
  )
```


Use fct_reorder to reorder the categorical variable(state) according to the order of percent (numeric)

```{r}
marj_use_df |> 
  filter(
    age == "12-17",
    #!State %in% c("Total U.S", "South") do this if you want to filter out all states except total US and South 
    ) |> 
  mutate(
    State = fct_reorder(State, percent)
  ) |> 
  ggplot(aes(x = State, y = percent, color = year))+
  geom_point()+
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

## NYC Restaurant Inspections

```{r}
data("rest_inspec") 

rest_inspec |> 
  count(boro, grade) |> 
  pivot_wider(
    names_from = grade,
    values_from = n
  )
  
```

```{r}
rest_inspec =
  rest_inspec |> 
  filter(
    str_detect(grade, "[A-C]"),
    !(boro == "Missing")
  )
```

Look at pizza places. dba variables 

```{r}
rest_inspec |> 
  mutate(dba = str_to_sentence(dba)) |> 
  filter(str_detect(dba, "Pizza"))
```

Visualization of boroughs that have the most pizza places. Use fct_infreq to order them according to frequency. Order matters in which str and fct function you use. 
```{r}
rest_inspec |> 
  mutate(dba = str_to_sentence(dba)) |> 
  filter(str_detect(dba, "Pizza")) |> 
  mutate(
    boro = fct_infreq(boro),
    boro = fct_recode(boro, "THE CITY" = "MANHATTAN")
  ) |> 
  ggplot(aes(x = boro)) +
  geom_bar()
```


One last thing on factors 
Regression coefficient. 

```{r}
rest_inspec |> 
  mutate(dba = str_to_sentence(dba)) |> 
  filter(str_detect(dba, "Pizza")) |> 
  mutate(boro = fct_infreq(boro)) |> 
  lm(zipcode ~ boro, data = _) 
  
```

```{r}

```

